---
title: "Project_2"
author: "Roger Dennison"
date: "2023-03-11"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#install.packages("devtools")
#install.packages("mlbench")
#install.packages("klaR")
#install.packages("htmltools")
## avoid converting warnings to errors by setting this environment variable to true (from within R)
Sys.setenv("R_REMOTES_NO_ERRORS_FROM_WARNINGS"=TRUE)
#library(devtools) # if not installed do install.package('devtools')library(remotes) # if not installed do install.package('remotes')
```


```{r}
#load the mlbench package which has the BreastCancer data set
require(mlbench)
```


```{r}
# if you don't have any required package, use the install.packages() command
# load the data set
data(BreastCancer)
ls(BreastCancer)
```


```{r}
# some algorithms don't like missing values, so remove rows with missing values
BreastCancer <- na.omit(BreastCancer)
```


```{r}
# remove the unique identifier, which is useless and would confuse the machine learning algorithms
BreastCancer$Id <- NULL 
head(BreastCancer)
str(BreastCancer)
df2 <- data.frame(sapply(BreastCancer[1:9], function(x) as.numeric(as.character(x))))
z <- scale(df2[,1:9],center=TRUE,scale=TRUE)
head(z)
```



```{r}
library(e1071)

mysvm <- svm(Class ~ ., BreastCancer)
mysvm.pred <- predict(mysvm, BreastCancer)
table(mysvm.pred,BreastCancer$Class)
```

Tried to include Naive Bayes model, but Klar devtools would not work no matter how many attempts I made. Proceeded without those models.

```{r}
#install.packages("klaR")
#install.packages("remotes")
#library(remotes)
#library(klaR)
#library(devtools)
#mynb <- NaiveBayes(Class ~ ., BreastCancer)
#mynb.pred <- predict(mynb,BreastCancer)
#head(mynb.pred$class)
#table(mynb.pred$class,BreastCancer$Class)
#str(mysvm.pred)
#str(mynb.pred)
#              setosa versicolor virginica
#   setosa     50      0          0
#   versicolor  0     47          3
#   virginica   0      3         47
```



```{r}
library(nnet)
library(neuralnet)
str(BreastCancer)
for (i in c(1:9)){
BreastCancer[,i] <-(as.numeric(BreastCancer[,i])-min(as.numeric(BreastCancer[,i]))) /
  (max(as.numeric(BreastCancer[,i]))-min(as.numeric(BreastCancer[,i])))
}
mynnet <- neuralnet(Class ~ ., BreastCancer, hidden=c(5,4))
head(BreastCancer)
str(mynnet)
library(neuralnet)
mynnet.pred <- neuralnet::compute(mynnet, BreastCancer) #Get the actual classes out
predicted.class <- apply(mynnet.pred$net.result,1,which.max)-1
mynnet.pred <- predicted.class
```


```{r}
library(MASS)

#Decision trees
library(rpart)
mytree <- rpart(Class ~ ., BreastCancer)
plot(mytree); text(mytree) 
summary(mytree)
mytree.pred <- predict(mytree,BreastCancer,type="class")
table(mytree.pred,BreastCancer$Class)
```


```{r}
# Leave-1-Out Cross Validation (LOOCV)
ans <- numeric(length(BreastCancer[,1]))
for (i in 1:length(BreastCancer[,1])) {
  mytree <- rpart(Class ~ ., BreastCancer[-i,])
  mytree.pred <- predict(mytree,BreastCancer[i,],type="class")
  ans[i] <- mytree.pred
}
ans <- factor(ans,labels=levels(BreastCancer$Class))
table(ans,BreastCancer$Class)
```



```{r}
# The same as above in this case


#Quadratic Discriminant Analysis
library(MASS)

myqda <- qda(Class ~ ., BreastCancer)
myqda.pred <- predict(myqda, BreastCancer)
head(myqda.pred$class)
table(myqda.pred$class,BreastCancer$Class)
```


Tried to include rda model model, but Klar devtools would not work no matter how many attempts I made. Proceeded without that model.

```{r}
#Regularised Discriminant Analysis
#library(klaR)
#myrda <- rda(Class ~ ., BreastCancer)
#myrda.pred <- predict(myrda, BreastCancer)

#table(myrda.pred$class,BreastCancer$Class)
```



```{r}
#Random Forests
library(randomForest)
myrf <- randomForest(Class ~ ., BreastCancer)
myrf.pred <- predict(myrf, BreastCancer)
head(myrf.pred)
table(myrf.pred, BreastCancer$Class)
```

Tried to include all 6, but RDA and Naive Bayes would not work no matter how many attempts I made. Proceeded below with only 4 models. 


```{r}
combine.classes<-data.frame(myrf.pred,
                            mytree.pred,mynnet.pred,mysvm.pred)
```

```{r}
#I wanted to get the best prediction, so I chose to only trust when 3 of 4 predictions agree.

head(combine.classes)
head(myrf.pred)

combine.classes$myrf.pred<-ifelse(combine.classes$myrf.pred=="benign", 0, 1)
combine.classes[,2]<-ifelse(combine.classes[,2]=="benign", 0, 1)
combine.classes[,3]<-ifelse(combine.classes[,3]=="benign", 0, 1)
combine.classes[,4]<-ifelse(combine.classes[,4]=="benign", 0, 1)

#Conduct Majority Vote
combine.classes$majority.vote<- rowSums(combine.classes)
combine.classes$class <-ifelse(combine.classes$majority.vote>=3, "malignant","benign")
table(BreastCancer$Class, combine.classes$class )

```

